### 2014-12-19 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: more updates to report

\chapter*{\centering Abstract}
\pagenumbering{gobble}

@@

\chapter{Fire Incident Summary}
\label{fire_sum}
The account of events for this incident was documented by the Safety Investigation Team of the Prince George's County Fire/EMS Department~\cite{PGCounty2013}. At 21:11 hours on February 24, 2012 Prince George's Fire/EMS Department responded to a structure fire in Riverdale Heights, Maryland with a fire in the basement~\cite{PGCounty2013}. For the purpose of this analysis, the details regarding the timeline are considered to be approximate values. The Prince George's County report~\cite{PGCounty2013} includes narratives of actions of the response personnel. The following narrative of the incident, focused on the events which led to burn injuries to two emergency-personnel, was taken from the Prince George's County report~\cite{PGCounty2013}:
\begin{quote}
Truck 809 responded from quarters, as the first due truck company, and arrived on 57th Avenue right behind Engine 807B. When Truck 809 stopped to let Engine 807B layout, Truck 809 Can dismounted the truck and had to walk up the street as it pulled away. Truck 809 Officer observed a lot of smoke moving extremely fast through the front yard and across the street upon exiting the truck. Truck 809 Officer and Truck 809 Forcible Entry proceeded directly to the front door \ldots of the structure. Both of them entered the structure with full PPE, including SCBA, but without the protection of a hose line. Truck 809 Officer and Truck 809 Forcible Entry began primary searches on the first floor. At some point during the primary search, as conditions worsened, the front door closed, trapping Truck 809 Officer and Truck 809 Forcible Entry inside the structure. This situation could not have occurred if a charged hose line was operating inside the structure at the time the door slammed shut. They were the only firefighters operating on the first floor at that time.

@@

\chapter{Summary}

@@

>>>>

\chapter*{\centering Abstract}
Fire Dynamics Simulator (FDS), which is a fire model that is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on February 24, 2012, within a single-story plus basement, single-family residential structure in Riverdale Heights, MD, that resulted in the injury of two firefighters. The inputs for the FDS simulation are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. After the front door was opened, the interior stairwell, aided by strong winds blowing into the basement, acted as a chimney for hot gases in the basement to flow towards regions of lower pressure, out through the front door. The temperature of the gases in the interior stairwell was estimated to be in excess of 400~$^{\circ}$C (750~$^{\circ}$F). Two firefighters were located in the flow path between the interior stairwell door and the front door of the structure, were exposed to these elevated temperatures, and suffered serious burn injures. Four additional firefighters were injured during fire operations on the first floor.
\pagenumbering{gobble}

@@

\chapter{Fire Incident Summary}
\label{fire_sum}
The account of events for this incident was documented by the Safety Investigation Team of the Prince George's County Fire/EMS Department~\cite{PGCounty2013}. At 21:11 hours on February 24, 2012 Prince George's Fire/EMS Department responded to a structure fire in Riverdale Heights, Maryland with a fire in the basement~\cite{PGCounty2013}. For the purpose of this analysis, the details regarding the timeline are considered to be approximate values. The Prince George's County report~\cite{PGCounty2013} includes narratives of actions of the response personnel. The following narrative of the incident, focused on the events which led to serious burn injuries to two emergency-personnel, was taken from the Prince George's County report~\cite{PGCounty2013}:
\begin{quote}
Truck 809 responded from quarters, as the first due truck company, and arrived on 57th Avenue right behind Engine 807B. When Truck 809 stopped to let Engine 807B layout, Truck 809 Can dismounted the truck and had to walk up the street as it pulled away. Truck 809 Officer observed a lot of smoke moving extremely fast through the front yard and across the street upon exiting the truck. Truck 809 Officer and Truck 809 Forcible Entry proceeded directly to the front door \ldots of the structure. Both of them entered the structure with full PPE, including SCBA, but without the protection of a hose line. Truck 809 Officer and Truck 809 Forcible Entry began primary searches on the first floor. At some point during the primary search, as conditions worsened, the front door closed, trapping Truck 809 Officer and Truck 809 Forcible Entry inside the structure. This situation could not have occurred if a charged hose line was operating inside the structure at the time the door slammed shut. They were the only firefighters operating on the first floor at that time.

@@

\chapter{Summary}
Fire Dynamics Simulator was used to provide insight into the fire dynamics of a fire that occurred within a single-story plus basement, single-family residential structure in Riverdale Heights, MD, that resulted in the serious injury of two firefighters. The fuel, fire size, and fire growth rate that were used in the FDS simulation were estimated by taking into account all of the available information including the post incident report from Prince George's Country Fire/EMS~\cite{PGCounty2013}, post-incident pictures, and relevant literature. This resulted in a maximum specified source fire of approximately 8.5~MW in the basement. Based on the limited ventilation conditions in the basement, the FDS simulation results indicated that before the front door was opened, there was combustion occurring outside the structure at the rear basement windows which had failed prior to arrival. Once the front door was opened, there was sufficient oxygen for combustion to occur within the structure, which intermittent flaming combustion on the first floor.

The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards regions of lower pressure through the open front door of the structure. After the front door was opened, a flow path was established between the basement and the front side of the structure (the front door). The opening of the front door resulted in a rapid change in the conditions within the flow path. In the interior stairwell, the flow velocities were greater than 9~m/s (20~mph) and the temperature of the gases was estimated to be in excess of 400~$^{\circ}$C (750~$^{\circ}$F), which exceeds the Class III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F).

Two firefighters were located in the flow path between the basement and the doors on the front side of the structure. One firefighter was able to break a window to escape the hazardous conditions and after a call for assistance the second firefighter was removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where they were treated for their injuries. Four other firefighters were injured on-scene and transported to the hospital for treatment.

@@

The authors would like to thank XXXXX of the Prince George's County Fire/EMS Department for their assistance with this study and their dedication to improving firefighter safety. The authors would also like to thank John Culbertson, Director of the Montana Fire Services Training School, for his assistance in researching and compiling the flow path related LODD/LODI incidents shown in Table~\ref{tab:LODD}. Finally, the authors would like to thank Joseph Willi of the University of Illinois Urbana-Champaign for his assistance in creating dimensioned drawings of the structure.
